KidsAIR initial SUMs data work

### 2020-03-18 exposure_initial_analysis.Rmd
exposure analysis updates

pm_home_char_results <- lm(mean_pm ~ home_insert, data = pm_home_char)
summary(pm_home_char_results)
plot(pm_home_char_results)
```

@@

pm_stove_use_results <- lm(mean_pm ~ wood_cords_cat, data = pm_stove_use)
summary(pm_stove_use_results)
plot(pm_stove_use_results)
pm_stove_use_plot <- pm_stove_use %>% 
  ggplot() + 
    geom_boxplot(aes(burn_level, mean_pm))
pm_stove_use_plot
```

@@

pm_stove_grades_results <- lm(mean_pm ~ stove_grade, data = pm_stove_grades)
summary(pm_stove_grades_results)
anova(pm_stove_grades_results)
plot(pm_stove_grades_results)
pm_stove_grades_plot <- pm_stove_grades %>% 
  ggplot() + 
    geom_boxplot(aes(stove_grade, mean_pm))
pm_stove_grades_plot
```

@@

pm_home_act_results <- lmer(mean_pm ~ home_act_sweep + (1 | home), data = pm_home_act)
summary(pm_home_act_results)
plot(pm_home_act_results)
```

@@

pm_moisture_results <- lm(mean_pm ~ moisture_ave_5, data = pm_moisture)
summary(pm_moisture_results)
plot(pm_moisture_results)
```

@@

pm_home_act_results <- lm(mean_pm ~ rh_min_cat, data = pm_temp_rh)
summary(pm_home_act_results)
plot(pm_home_act_results)
```

>>>>

pm_home_char_results <- lm(log(mean_pm) ~ home_insert, data = pm_home_char)
summary(pm_home_char_results)
plot(pm_home_char_results)
```

@@

pm_stove_use_results <- lm(log(mean_pm) ~ wood_collect_method, data = pm_stove_use)
summary(pm_stove_use_results)
plot(pm_stove_use_results)
pm_stove_use_plot <- pm_stove_use %>% 
  ggplot() + 
    geom_boxplot(aes(burn_level, mean_pm))
pm_stove_use_plot
```

@@

pm_stove_grades_results <- lm(log(mean_pm) ~ stove_grade, data = pm_stove_grades)
summary(pm_stove_grades_results)
anova(pm_stove_grades_results)
plot(pm_stove_grades_results)
pm_stove_grades_plot <- pm_stove_grades %>% 
  ggplot() + 
    geom_boxplot(aes(stove_grade, mean_pm))
pm_stove_grades_plot
```

@@

pm_home_act_results <- lmer(log(mean_pm) ~ home_act_sweep + (1 | home), data = pm_home_act)
summary(pm_home_act_results)
plot(pm_home_act_results, 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(pm_home_act_results), main = "QQ Plot")
```

@@

pm_moisture_results <- lm(log(mean_pm) ~ moisture_cat, data = pm_moisture)
summary(pm_moisture_results)
plot(pm_moisture_results)
```

@@

pm_home_act_results <- lm(log(mean_pm) ~ rh_min_cat, data = pm_temp_rh)
summary(pm_home_act_results)
plot(pm_home_act_results)
```

# SUMs 
```{r}
#### Join PM dataset with full SUMs dataset

pm_join <- read_rds("Output/pm_clean.rds") %>% 
  mutate(sample_datetime = ymd_hms(pm_datetime_new)) %>% 
  separate(sample_datetime, c("sample_date", "sample_time"), sep = " ", remove = FALSE) %>% 
  group_by(area, home) %>% 
  arrange(sample_datetime) %>% 
  mutate(pm_last_datetime = last(sample_datetime),
         pm_first_datetime = first_datetime_new,
         sample_datetime = round_date(sample_datetime, unit = "minute"),
         pm_rolling_20 = rollmean(pm, 20, fill = "extend")) %>% 
  ungroup() %>% 
  arrange(area, home, sample_datetime)

sums_join <- read_rds("Output/sums_clean.rds") %>% 
  mutate(sample_datetime = ymd_hms(datetime_sums)) %>% 
  separate(sample_datetime, c("sample_date", "sample_time"), sep = " ", remove = FALSE) %>% 
  mutate(sample_datetime = round_date(sample_datetime, unit = "minute")) %>% 
  ungroup() %>% 
  arrange(area, home, sample_datetime)

pm_sums_joined <- sums_join %>% 
  left_join(pm_join, by = c("area", "home", "sample_datetime")) %>% 
  filter(pm_rolling_20 < 1000) %>% 
  filter(!is.na(temp_c_sums)) %>% 
  filter(!is.na(pm_rolling_20)) %>% 
  mutate(temp_c_sums_5 = temp_c_sums/5)

pm_sums_count <- pm_sums_joined %>% 
  count(home)

cor.test(pm_sums_joined$temp_c_sums, pm_sums_joined$pm_rolling_20, method = "spearman")
pm_sums_model <- lmer(log(pm_rolling_20) ~ temp_c_sums_5 + (1 | home), data = pm_sums_joined)

summary(pm_sums_model)
exp((0.146291)-1)*100
plot(pm_sums_model, 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(pm_sums_model), main = "QQ Plot")

pm_sums_plot <- pm_sums_joined %>% 
  ggplot() + 
    geom_point(aes(log(pm_rolling_20), temp_c_sums)) +
    geom_smooth(aes(log(pm_rolling_20), temp_c_sums)) +
    theme_minimal()
pm_sums_plot

#### Join PM dataset with SUMs events datasets

# Mean PM for each placebo home over winter 1
pm_home_mean_date <- pm_clean %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  mutate(pm = if_else(pm > 500, 500, pm)) %>% 
  separate(pm_datetime_new, c("date_sums", "pm_time"), sep = " ", remove = TRUE) %>% 
  group_by(area, home, date_sums) %>% 
  summarize(mean_pm = mean(pm, na.rm = TRUE)) %>% 
  group_by(area, home, date_sums) %>% 
  distinct(date_sums, .keep_all = TRUE) %>% 
  ungroup() %>% 
  arrange(area, home, date_sums)

pm_sums_events_joined <- pm_home_mean_date %>% 
  left_join(sums_events_27, by = c("area", "home", "date_sums")) %>% 
  filter(sums_event == 1) %>% 
  arrange(area, home, date_sums) %>% 
  group_by(area, home, mean_pm) %>% 
  summarize(sums_events = n()) %>% 
  #mutate(sums_events = factor(sums_events))
  mutate(sums_events_cat = cut(sums_events, breaks = c(0, 1, 10),
                           labels = c("1", ">1")))

median(pm_sums_events_joined$sums_events)

pm_sums_model <- lmer(mean_pm ~ sums_events_cat + (1 | home), data = pm_sums_events_joined)
pm_sums_model <- lmer(log(mean_pm) ~ sums_events_cat + (1 | home), data = pm_sums_events_joined)

summary(pm_sums_model)
#exp((0.146291)-1)*100
plot(pm_sums_model, 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(pm_sums_model), main = "QQ Plot")
```

### 2020-03-18 sums_data_cleaning.Rmd
exposure analysis updates

thresholds_test <- sums_clean %>% 
  mutate(datetime_sums = ymd_hms(datetime_sums)) %>% 
  group_by(area, home, winter_id) %>%
  mutate(threshold_temp = if_else(temp_c_sums >= 40, 1, 0),
         threshold_temp_event = if_else(threshold_temp == 1 & lag(threshold_temp) == 0, 
                                        "start", "Temp < threshold"),
         threshold_temp_event = if_else(is.na(threshold_temp_event), "start", threshold_temp_event),
         threshold_temp_event = if_else(threshold_temp == 1 & threshold_temp_event == "Temp < threshold",
                                      "Temp >= threshold", threshold_temp_event),
         threshold_temp_event = if_else(threshold_temp_event == "Temp >= threshold" &
                                      lead(threshold_temp_event) == "Temp < threshold", 
                                      "event", threshold_temp_event)) %>% 
  filter(threshold_temp_event == "event" | threshold_temp_event == "start") %>% 
  spread(threshold_temp_event, datetime_sums) %>% 
  mutate(event = lead(event),
         time_diff_min = (event - start),
         time_diff_min = as.numeric(time_diff_min/60)) %>% 
  filter(!is.na(time_diff_min)) %>% 
  mutate(next_event = (lead(start)-event),
         next_event = as.numeric(next_event)/3600,
         event2 = if_else(next_event <= 1.5, lead(event), event),
         time_diff_min2 = (event - start),
         time_diff_min2 = as.numeric(time_diff_min2/60),
         event3 = if_else(next_event <= 1.5 & lead(next_event) <= 1.5, lead(event, 2), event2),
         time_diff_min3 = (event3 - start),
         time_diff_min3 = as.numeric(time_diff_min3/60)) %>% 
  distinct(event3, .keep_all = TRUE) %>% 
  mutate(time_diff_min4 = (event3 - start),
         time_diff_min4 = as.numeric(time_diff_min4/60)) %>% 
  filter(!is.na(time_diff_min4)) %>% 
  group_by(area, home, winter_id, date_sums) %>% 
  summarize(n = n(), mean_time = mean(time_diff_min4)) %>% 
  group_by(area) %>% 
  summarize("Mean events per day" = round(mean(n), digits = 2),
            "Mean event length (min)" = round(mean(mean_time), digits = 0))
thresholds_test

# Function for calculating heating events using specified threshold temps
# Also specify the "gap" time between events
# Try another function for % day above threshold
thresholds_function <- function(data, temp, gap) {

thresholds <- data %>% 
  mutate(datetime_sums = ymd_hms(datetime_sums)) %>% 
  group_by(area, home, winter_id) %>%
  mutate(threshold_temp = if_else(temp_c_sums >= temp, 1, 0),
         threshold_temp_event = if_else(threshold_temp == 1 & lag(threshold_temp) == 0, 
                                        "start", "Temp < threshold"),
         threshold_temp_event = if_else(is.na(threshold_temp_event), "start", threshold_temp_event),
         threshold_temp_event = if_else(threshold_temp == 1 & threshold_temp_event == "Temp < threshold",
                                      "Temp >= threshold", threshold_temp_event),
         threshold_temp_event = if_else(threshold_temp_event == "Temp >= threshold" &
                                      lead(threshold_temp_event) == "Temp < threshold", 
                                      "event", threshold_temp_event)) %>% 
  filter(threshold_temp_event == "event" | threshold_temp_event == "start") %>% 
  spread(threshold_temp_event, datetime_sums) %>% 
  mutate(event = lead(event),
         time_diff_min = (event - start),
         time_diff_min = as.numeric(time_diff_min/60)) %>% 
  filter(!is.na(time_diff_min)) %>% 
  mutate(next_event = (lead(start)-event),
         next_event = as.numeric(next_event)/3600,
         event2 = if_else(next_event <= gap, lead(event), event),
         time_diff_min2 = (event - start),
         time_diff_min2 = as.numeric(time_diff_min2/60),
         event3 = if_else(next_event <= gap & lead(next_event) <= gap, lead(event, 2), event2),
         time_diff_min3 = (event3 - start),
         time_diff_min3 = as.numeric(time_diff_min3/60)) %>% 
  distinct(event3, .keep_all = TRUE) %>% 
  mutate(time_diff_min4 = (event3 - start),
         time_diff_min4 = as.numeric(time_diff_min4/60)) %>% 
  filter(!is.na(time_diff_min4)) %>% 
  group_by(area, home, winter_id, date_sums) %>% 
  summarize(n = n(), mean_time = mean(time_diff_min4)) %>% 
  group_by(area) %>% 
  summarize("Mean events per day" = round(mean(n), digits = 2),
            "Mean event length (min)" = round(mean(mean_time), digits = 0))
thresholds

}

thresholds_function(sums_clean, 40, 2)

# Function to look at percentage of time spent above a threshold
thresholds_percent <- function(data, temp) {

thresholds <- data %>% 
  group_by(area, home, winter_id, date_sums) %>%
  mutate(threshold_temp = if_else(temp_c_sums >= temp, 1, 0)) %>% 
  summarize(percent_above = sum(threshold_temp)/n()*100) %>% 
  group_by(area) %>% 
  summarize("% time/day > threshold" = round(mean(percent_above), digits = 2))
thresholds

}

thresholds_percent(sums_clean, 40)
```

>>>>

sums_events_25 <- sums_clean %>% 
  mutate(datetime_sums = ymd_hms(datetime_sums)) %>% 
  group_by(area, home, winter_id) %>%
  mutate(threshold_temp = if_else(temp_c_sums >= 25, 1, 0),
         threshold_temp_event = if_else(threshold_temp == 1 & lag(threshold_temp) == 0, 
                                        "start", "Temp < threshold"),
         threshold_temp_event = if_else(is.na(threshold_temp_event), "start", threshold_temp_event),
         threshold_temp_event = if_else(threshold_temp == 1 & threshold_temp_event == "Temp < threshold",
                                      "Temp >= threshold", threshold_temp_event),
         threshold_temp_event = if_else(threshold_temp_event == "Temp >= threshold" &
                                      lead(threshold_temp_event) == "Temp < threshold", 
                                      "event_end", threshold_temp_event)) %>% 
  filter(threshold_temp_event == "event_end" | threshold_temp_event == "start") %>% 
  mutate(time_sums = as.character(time_sums),
         date_sums = as.character(date_sums)) %>%  
  spread(threshold_temp_event, datetime_sums) %>% 
  mutate(event_end = lead(event_end)) %>% 
  filter(!is.na(start)) %>% 
  mutate(hrs_till_next_event = (lead(start)-event_end),
         hrs_till_next_event = as.numeric(hrs_till_next_event)/3600,
         event_end2 = if_else(hrs_till_next_event <= 1.5, lead(event_end), event_end),
         event_end3 = if_else(hrs_till_next_event <= 1.5 & lead(hrs_till_next_event) <= 1.5, 
                              lead(event_end, 2), event_end2),
         event_end4 = if_else(hrs_till_next_event <= 1.5 & lead(hrs_till_next_event) <= 1.5
                              & lead(hrs_till_next_event, 2) <= 1.5, 
                              lead(event_end, 3), event_end3),
         event_end5 = if_else(hrs_till_next_event <= 1.5 & lead(hrs_till_next_event) <= 1.5
                              & lead(hrs_till_next_event, 2) <= 1.5 & lead(hrs_till_next_event, 3) <= 1.5, 
                              lead(event_end, 4), event_end4),
         event_end6 = if_else(hrs_till_next_event <= 1.5 & lead(hrs_till_next_event) <= 1.5
                              & lead(hrs_till_next_event, 2) <= 1.5 & lead(hrs_till_next_event, 3) <= 1.5
                              & lead(hrs_till_next_event, 4) <= 1.5, 
                              lead(event_end, 5), event_end5)) %>% 
  distinct(event_end6, .keep_all = TRUE) %>% 
  mutate(event_length_min = (event_end6 - start),
         event_length_min = as.numeric(event_length_min/60),
         hrs_till_next_event = (lead(start)-event_end6),
         hrs_till_next_event = as.numeric(hrs_till_next_event)/3600) %>% 
  filter(!is.na(event_length_min)) %>% 
  group_by(area, home, winter_id, date_sums) %>% 
  mutate(event_start_datetime = start,
         event_end_datetime = event_end6,
         sums_event = as.numeric(1)) %>% 
  select(area, home, home_winter_id, winter_id, date_sums, 
         event_start_datetime, event_end_datetime, sums_event,
         event_length_min, hrs_till_next_event)

sums_events_27 <- sums_clean %>% 
  mutate(datetime_sums = ymd_hms(datetime_sums)) %>% 
  group_by(area, home, winter_id) %>%
  mutate(threshold_temp = if_else(temp_c_sums >= 27, 1, 0),
         threshold_temp_event = if_else(threshold_temp == 1 & lag(threshold_temp) == 0, 
                                        "start", "Temp < threshold"),
         threshold_temp_event = if_else(is.na(threshold_temp_event), "start", threshold_temp_event),
         threshold_temp_event = if_else(threshold_temp == 1 & threshold_temp_event == "Temp < threshold",
                                      "Temp >= threshold", threshold_temp_event),
         threshold_temp_event = if_else(threshold_temp_event == "Temp >= threshold" &
                                      lead(threshold_temp_event) == "Temp < threshold", 
                                      "event_end", threshold_temp_event)) %>% 
  filter(threshold_temp_event == "event_end" | threshold_temp_event == "start") %>% 
  mutate(time_sums = as.character(time_sums),
         date_sums = as.character(date_sums)) %>%  
  spread(threshold_temp_event, datetime_sums) %>% 
  mutate(event_end = lead(event_end)) %>% 
  filter(!is.na(start)) %>% 
  mutate(hrs_till_next_event = (lead(start)-event_end),
         hrs_till_next_event = as.numeric(hrs_till_next_event)/3600,
         event_end2 = if_else(hrs_till_next_event <= 1.5, lead(event_end), event_end),
         event_end3 = if_else(hrs_till_next_event <= 1.5 & lead(hrs_till_next_event) <= 1.5, 
                              lead(event_end, 2), event_end2),
         event_end4 = if_else(hrs_till_next_event <= 1.5 & lead(hrs_till_next_event) <= 1.5
                              & lead(hrs_till_next_event, 2) <= 1.5, 
                              lead(event_end, 3), event_end3),
         event_end5 = if_else(hrs_till_next_event <= 1.5 & lead(hrs_till_next_event) <= 1.5
                              & lead(hrs_till_next_event, 2) <= 1.5 & lead(hrs_till_next_event, 3) <= 1.5, 
                              lead(event_end, 4), event_end4),
         event_end6 = if_else(hrs_till_next_event <= 1.5 & lead(hrs_till_next_event) <= 1.5
                              & lead(hrs_till_next_event, 2) <= 1.5 & lead(hrs_till_next_event, 3) <= 1.5
                              & lead(hrs_till_next_event, 4) <= 1.5, 
                              lead(event_end, 5), event_end5)) %>% 
  distinct(event_end6, .keep_all = TRUE) %>% 
  mutate(event_length_min = (event_end6 - start),
         event_length_min = as.numeric(event_length_min/60),
         hrs_till_next_event = (lead(start)-event_end6),
         hrs_till_next_event = as.numeric(hrs_till_next_event)/3600) %>% 
  filter(!is.na(event_length_min)) %>% 
  group_by(area, home, winter_id, date_sums) %>% 
  mutate(event_start_datetime = start,
         event_end_datetime = event_end6,
         sums_event = as.numeric(1)) %>% 
  select(area, home, home_winter_id, winter_id, date_sums, 
         event_start_datetime, event_end_datetime, sums_event,
         event_length_min, hrs_till_next_event)

sums_events_30 <- sums_clean %>% 
  mutate(datetime_sums = ymd_hms(datetime_sums)) %>% 
  group_by(area, home, winter_id) %>%
  mutate(threshold_temp = if_else(temp_c_sums >= 30, 1, 0),
         threshold_temp_event = if_else(threshold_temp == 1 & lag(threshold_temp) == 0, 
                                        "start", "Temp < threshold"),
         threshold_temp_event = if_else(is.na(threshold_temp_event), "start", threshold_temp_event),
         threshold_temp_event = if_else(threshold_temp == 1 & threshold_temp_event == "Temp < threshold",
                                      "Temp >= threshold", threshold_temp_event),
         threshold_temp_event = if_else(threshold_temp_event == "Temp >= threshold" &
                                      lead(threshold_temp_event) == "Temp < threshold", 
                                      "event_end", threshold_temp_event)) %>% 
  filter(threshold_temp_event == "event_end" | threshold_temp_event == "start") %>% 
  mutate(time_sums = as.character(time_sums),
         date_sums = as.character(date_sums)) %>%  
  spread(threshold_temp_event, datetime_sums) %>% 
  mutate(event_end = lead(event_end)) %>% 
  filter(!is.na(start)) %>% 
  mutate(hrs_till_next_event = (lead(start)-event_end),
         hrs_till_next_event = as.numeric(hrs_till_next_event)/3600,
         event_end2 = if_else(hrs_till_next_event <= 1.5, lead(event_end), event_end),
         event_end3 = if_else(hrs_till_next_event <= 1.5 & lead(hrs_till_next_event) <= 1.5, 
                              lead(event_end, 2), event_end2),
         event_end4 = if_else(hrs_till_next_event <= 1.5 & lead(hrs_till_next_event) <= 1.5
                              & lead(hrs_till_next_event, 2) <= 1.5, 
                              lead(event_end, 3), event_end3),
         event_end5 = if_else(hrs_till_next_event <= 1.5 & lead(hrs_till_next_event) <= 1.5
                              & lead(hrs_till_next_event, 2) <= 1.5 & lead(hrs_till_next_event, 3) <= 1.5, 
                              lead(event_end, 4), event_end4),
         event_end6 = if_else(hrs_till_next_event <= 1.5 & lead(hrs_till_next_event) <= 1.5
                              & lead(hrs_till_next_event, 2) <= 1.5 & lead(hrs_till_next_event, 3) <= 1.5
                              & lead(hrs_till_next_event, 4) <= 1.5, 
                              lead(event_end, 5), event_end5)) %>% 
  distinct(event_end6, .keep_all = TRUE) %>% 
  mutate(event_length_min = (event_end6 - start),
         event_length_min = as.numeric(event_length_min/60),
         hrs_till_next_event = (lead(start)-event_end6),
         hrs_till_next_event = as.numeric(hrs_till_next_event)/3600) %>% 
  filter(!is.na(event_length_min)) %>% 
  group_by(area, home, winter_id, date_sums) %>% 
  mutate(event_start_datetime = start,
         event_end_datetime = event_end6,
         sums_event = as.numeric(1)) %>% 
  select(area, home, home_winter_id, winter_id, date_sums, 
         event_start_datetime, event_end_datetime, sums_event,
         event_length_min, hrs_till_next_event)

#write_rds(sums_events_25, "Output/sums_events_25.rds")
#write_rds(sums_events_27, "Output/sums_events_27.rds")
#write_rds(sums_events_30, "Output/sums_events_30.rds")
```
